---
title: "MEMgene paper analysis"
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_document:
    theme: flatly
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: no
---

Analysis of simulated datasets from Lotterhos & Whitlock with the same random sampling of 90 sites as in Wagner et al. (2017).

## 1. Preparations

### a) Packages

```{r packages}
library(hierfstat)
library(memgene)
library(here)
library(parallel)
library(dplyr)
library(spdep)
```

### b) Import spatial coordinates

```{r coords}
Coords <- list()
Coords$Pairs <- list()
Coords$Transect <- list()
Coords$Random <- list()
Coords$Random$E453 <- Coords$Random$E988 <- Coords$Random$E950 <- 
  read.table("SchemeRandom1.txt") # upload this file

for(i in 1:length(Coords))
{
  if(length(Coords[[i]]) > 0)
  {
    for(k in 1:length(Coords[[i]]))
    {
      b <- order(Coords[[i]][[k]][,3], Coords[[i]][[k]][,2]) # Sort by y, then x
      Coords[[i]][[k]] <- Coords[[i]][[k]][b,]  # correct order!!
    }
  }
}
```

### c) Parameter space

Genetic data: find all file names in the folder "SimFilesLFMM" that contain 'lfmm':

```{r files}
Filenames.lfmm <- list.files(paste0(here::here(), "/dryad//SimFilesLFMM/"), pattern="lfmm") 

test <- Reduce(rbind, strsplit(Filenames.lfmm, split="[_=]"))
test <- cbind(test, Reduce(rbind, strsplit(test[,ncol(test)], split="[.]")))
tmp <- strsplit(test[,2], split="[.xs]")
test2 <- matrix(NA, nrow(test), 3)
for(i in 1:nrow(test)) test2[i,1:length(tmp[[i]])] <- tmp[[i]]
test2 <- gsub("[A-Z, a-z]","", test2)
test <- cbind(test, test2)
dimnames(test) <- list(NULL, c("Demography", "Design", "ID", "Env", "ID2", 
                               "V6", "NumPops", "V8", "V9", "NumInd", "Design2",
                               "NumPops2", "NumTrans", "NumInd2"))
```

Design matrix (parameter space): each row is one combination of parameter settings:

- Demography: single refugium (1R), two refugia (2R), isolation by distane (IBD), island model (IM)
- Design: sampling design (R, P, T; see below) and number of pops sampled
- Env: which of the three replicate landscapes '453', '950', '988'
- NumPops: how many populations are sampled
- NumInd: how many individuals sampled per pop
- Type: random (R), pairs (P), transects (T)

```{r Design}   
   Design <- as.data.frame(test[,c(1,2,4,7,10,13,14)])
   Design$Env <- ordered(Design$Env, levels=c(453,988,950))
   Design$Type <- ordered(substr(Design$Design, 1, 1), levels=c("P", "T", "R"))
   Design$Design <- as.character(Design$Design)
   Design$Design[Design$Design == "T30.T3x10"] <- "T30.3x10s"
   Design$Design[Design$Design == "T30.T6x5s"] <- "T30.6x5s"
   Design$NumPops <- as.numeric(as.character(Design$NumPops))
   head(Design)
```

### d) Create R replicate random samples of n = 30 sites

```{r subsets}
R = 10
Sites.30 <- list()
set.seed(19)
for(r in 1:R)
{
  Sites.30[[r]] <- sort(sample(1:90, 30, replace=FALSE))
}

saveRDS(Sites.30, paste0(here::here(), "/output/Sites.30.rds"))
```

### e) Create R replicate random samples of n = 60 sites

```{r}
R = 10
Results.drop <- rep( list(list()), R)

set.seed(297)
Drop <-lapply(c(1:R), function(r) sort(sample(1:90, 30, replace=FALSE)))

saveRDS(Drop, paste0(here::here(), "/output/Drop.rds"))
```


## 2. Genetic data

### a) Select datasets

Select all data files with the largest sample size (90) and random sampling (R)

```{r}
Sites.R.90 <- c(1:nrow(Design))[Design$NumPops==90 & Design$Type=="R"] 
```

Check parameters for selected datasets:

```{r}
Design[Sites.R.90,]
```

### b) Run function 'getGenData' to extract the genetic data

This function:

- extracts the genetic data, 
- selects the 9900 neutral loci, 
- randomizes the order of the neutral loci (allows sampling loci),
- adds a first column 'pop' with site as a factor (format for functions in the package `hierfstat`), and
- randomizes the alleles of each SNP within each population (allows sampling individuals). 

Define function:

```{r getGenData}
getGenData <- function(j)
{
  # Select the sites that need to be sampled for this run:
  i=Sites.R.90[j]
  cat("j:", j, ", i:", i, "\n")
  
  # Each file has NumPops x NumInd rows (sampled individuals) and up to 10000 columns (loci)
  tmp <- read.table(paste0(here::here(),"/dryad/SimFilesLFMM/", Filenames.lfmm[[i]]))
  
  # Drop non-neutral loci
  tmp <- tmp[,1:9900]
  
  # Randomize order of neutral loci
  tmp <- tmp[,sample(1:9900)]
    
  # Site: create vector of sites = pops (for each row = individual)
  Site <- rep(1:Design$NumPops[i],each=as.numeric(as.vector(Design$NumInd)[i]))
  
  #Data.hierfstat <- data.frame(pop=factor(Site), tmp)
  Data.hierfstat <- data.frame(pop=Site, tmp)
  
  # NEW Dec 2020: Randomize alleles within and pops (will allow subsampling individuals)
  Data.hierfstat <- Reduce(rbind, lapply(split(Data.hierfstat, Site), 
                                               function(p) data.frame(sapply(p , sample))))
  
  return(Data.hierfstat)
}
```

Run in parallel and save results:

```{r GenData90}
start_time <- Sys.time()

Index.j <- 1:length(Sites.R.90)

Data.R.90 <- mclapply(Index.j, function(j) getGenData(j),
                 mc.cores=detectCores())
names(Data.R.90) <- Sites.R.90

end_time <- Sys.time()
end_time - start_time
cat("This job took", (end_time - start_time)/length(Index.j), "per dataset.")

saveRDS(Data.R.90, paste0(here::here(), "/output/Data.R.90.rds"))
```

Results are stored in a list (one element per site j with 90 samples), with each element:

- Data.frame with first column "pop" (factor of site IDs) and 9900 columns of neutral loci

  
### c) Run function 'getDgen' to calculate sample Fst and genetic distances

Define function:

```{r getDgen}
#Data.R.90 <- readRDS(paste0(here::here(), "/output/Data.R.90.rds"))

getDgen <- function(j, Data=Data.R.90, dist.method = c("Fst", "Dch"), nLoci=9900, sim=FALSE)
{
  # Randomize pop if sim=TRUE
  if(sim==TRUE) {Data = lapply(Data, function(ls) data.frame(pop=sample(ls$pop), ls[,-1]))} 
  
  Fst <- basic.stats(Data[[j]][,1:(nLoci+1)])$overall

  Fst.30 <- t(sapply(1:length(Sites.30), function(s) 
    basic.stats(filter(Data[[j]][, 1:(nLoci+1)], is.element(pop, Sites.30[[s]])))$overall))
  
  D <- list()
  for(k in 1:length(dist.method))
  {
    D[[k]] <- genet.dist(Data[[j]][,1:(nLoci+1)], method = dist.method[k])
  }
  names(D) <- dist.method
  
  return(list(Fst=Fst, Fst.30 = Fst.30, Dgen=D))
}
```

NEW Dec 2020: run for different numbers of loci, create list object

```{r}
loci <- c(9900, 3300, 500, 100, 50)

Dgen.R.90 <- list()

for(s in length(loci):1)
{
  cat(s)
  Dgen.R.90[[s]] <- mclapply(Index.j, function(j) getDgen(j, Data=Data.R.90, nLoci=loci[s], 
                                                  sim=FALSE, dist.method=c("Fst", "Dch")),
                 mc.cores=detectCores())
  names(Dgen.R.90[[s]]) <- Sites.R.90
}
names(Dgen.R.90) <- paste0("L", loci)

saveRDS(Dgen.R.90, paste0(here::here(), "/output/Dgen.R.90.rds"))
```

Dgen.R.90 is a list, with one element per number of loci (e.g.: 9900, 3300, 500, 100, 50).
Each list element is again a list with one element per site j with 90 samples. Each of these contains the following:

- One list element per number of loci
  - One list element per dataset (site j with 90 samples)
    - Fst: vector with 'overall' statistics returned by `basic.stats` function
    - Fst.30: matrix with statistics for ten subsets of n = 30 sites
    - Dgen: list of pairwise genetic distance matrices
        - Fst (Fst)
        - Cavalli-Sforza and Edwards Chord distance (Dch)
    
    
### d) Create randomized data (for standardized effect size)

This takes long! Only run for 500 loci and for Fst distance method.

```{r}
R = 200

Dgen.sim <- list()

for(r in 1:R)
{
  cat("\n", r, ": ")
  Dgen <- list(list(), list(), list(), list(), list())
  names(Dgen) <- paste0("L", loci)
  #for(s in length(loci):1)
  for(s in c(3))
  {
    cat(s)
    Dgen[[s]] <- mclapply(Index.j, function(j) getDgen(j, Data=Data.R.90, nLoci=loci[s], 
                                                    sim=TRUE, dist.method=c("Fst")),
                   mc.cores=detectCores())
    names(Dgen[[s]]) <- Sites.R.90
  }
  Dgen.sim[[r]] <- Dgen
}

saveRDS(Dgen.sim, paste0(here::here(), "/output/Dgen.sim.rds"))
```

    
### e) Run function 'getMEMgene' to obtain RsqAdj and Moran's I

This function extracts the grid coordinates, performs MEMgene analysis with forward selection to obtain the adjusted Rsquared value. In addition, it derives the unadjusted Rsquared value using the same set of selected MEM eigenvectors. It also calculates Moran's I for the genetic data by deriving a scalogram S (Rsquared for each MEM vector, which sum to 1 across all MEM vectors) and multiplying it with the rescaled MEM eigenvectors (Moran's I for each vector). Finally, it returns the limits for Moran's I of the genetic data (min and max of Moran's I values of MEM vectors).

Argument 'subset' specifies a subset of sites to be used (meta-analysis mode, performing MEM with subset only). Argument 'drop' specifies which sites should be dropped (for comparative mode, keeping MEM of full dataset).

Define function:

```{r getMEMgene}
getMEMgene <- function(j, Dgen = Dgen.R.90[[1]], subset=NULL, drop=NULL)
{
  i=Sites.R.90[j]
  cat("j:", j, ", i:", i, "\n")
  
  # Extract the grid coordinates of the sampled sites
  coord <- data.matrix(Coords[[as.numeric(Design$Type[i])]]
                     [[as.numeric(Design$Env[i])]][,2:3])
  if(length(subset) > 0) {coord <- coord[subset,]}
  if(length(drop) > 0) { coord.drop <- coord[-drop,]}
  
  Res <- list()
  for(k in 1: length(Dgen[[j]]$Dgen))
  {
    
    Y <- as.matrix(Dgen[[j]]$Dgen[[k]])
    if(length(drop) > 0) {Y = Y[-drop, -drop]}
    if(length(subset) > 0) {Y <- Y[subset, subset]}
    
    # memGene
    
    MEM <- mgMEM(dist(coord))
    
    if(length(drop) > 0) {MEM$vectorsMEM <- MEM$vectorsMEM[-drop,]}

    Positive <- mgForward(Y, MEM$vectorsMEM[, MEM$valuesMEM > 0])

    RsqAdjPos = 0
    RsqPos = 0
    if(!is.na(Positive$selectedRsqAdj)) 
    { 
      RsqAdjPos = Positive$selectedRsqAdj
      # Get R square (unadjusted)
      MEM$analysis <- mgRDA(Y, MEM$vectorsMEM[, Positive$selectedMEM], full=TRUE)
      RsqPos <- sum(diag(MEM$analysis$pred)) / (sum(diag(MEM$analysis$pred)) +
                                               sum(diag(MEM$analysis$resid)))
    }
    
    if(length(drop) > 0)
    {
      MEM.drop <- mgMEM(dist(coord.drop))
      
      Positive <- mgForward(Y, MEM.drop$vectorsMEM[ , MEM.drop$valuesMEM > 0])
      RsqAdjPos.drop = 0
      if(!is.na(Positive$selectedRsqAdj)) { RsqAdjPos.drop = Positive$selectedRsqAdj}
    }

    # Centre distance matrix
    n <- nrow(Y)
    row.wt = rep(1, nrow(Y))
    col.wt = rep(1, ncol(Y))
    st <- sum(col.wt)
    sr <- sum(row.wt)
    row.wt <- row.wt/sr
    col.wt <- col.wt/st
    Y <- -0.5 * (Y * Y)
    row.mean <- apply(row.wt * Y, 2, sum)
    col.mean <- apply(col.wt * t(Y), 2, sum)
    col.mean <- col.mean - sum(row.mean * col.wt)
    Y <- sweep(Y, 2, row.mean)
    G <- t(sweep(t(Y), 2, col.mean))
    
    # Get Rsq for each MEM vector from a separate dbRDA for each vector m
    X <- MEM$vectorsMEM
  
    S <- rep(0, ncol(X))
    for(m in 1:ncol(X))
    {
      if(var(X[,m]) > 0)
      {
        p = 1
        H <- X[,m] %*% solve(t(X[,m]) %*% X[,m]) %*% t(X[,m])
        I <- diag(n)
        res <- (I - H) %*% G %*% (I - H)
        S[m] <- 1 - sum(diag(res))/sum(diag(G))
      }
    }
    
    S.tot = sum(S)
    S.min = min(S)
    S[S < 0] <- 0
    
    Values <- MEM$valuesMEM / abs(sum(MEM$valuesMEM))
    Range <- range(Values)
    Morans.I <- as.vector(S %*% Values)
    
    Res[[k]] <- list(RsqAdjPos=RsqAdjPos, RsqPos=RsqPos, Morans.I=Morans.I, Range=Range)
  }
  names(Res) <- names(Dgen[[1]]$Dgen)
  
  return(Res)
}
```

Run across all levels of number of loci

```{r}
#Dgen.R.90 <- readRDS(paste0(here::here(), "/output/Dgen.R.90.rds"))

start_time <- Sys.time() 

Index.j <- 1:length(Sites.R.90)

Results.R.90 <- list()
for(s in length(loci):1)
{
  cat(s)
  Results.R.90[[s]] <- mclapply(Index.j, function(j) 
    getMEMgene(j, Dgen = Dgen.R.90[[s]], subset=NULL),
    mc.cores=detectCores())
  names(Results.R.90[[s]]) <- Sites.R.90
}
names(Results.R.90) <- paste0("L", loci)
  
end_time <- Sys.time()
end_time - start_time
cat("This job took", (end_time - start_time)/length(Index.j)/length(loci), "per dataset.")

saveRDS(Results.R.90, paste0(here::here(), "/output/Results.R.90.rds"))
```

Results.R.90 is a list, with one element per number of loci (e.g.: 9900, 3300, 500, 100, 50).
Each list element is again a list with one element per site j with 90 samples. Each of these contains the following:

- One list element per number of loci
  - One list element per dataset (site j with 90 samples)
    - One list element per genetic distance measure
      - RsqAdjPos: adjusted Rsquare from memgene, based on MEM with positive eigenvalues
      - Morans.I: Moran's I for the genetic data
      - Range: range (minimum and maximum) of Moran's I of MEM vectors (limits for Moran's I of genetic data)
      
      
      
Run for simulated data (only for s = 3, i.e., 500 loci)

```{r}
#Dgen.sim <- readRDS(paste0(here::here(), "/output/Dgen.sim.rds"))

start_time <- Sys.time() 

Index.j <- 1:length(Sites.R.90)

Results.sim <- list()
#for(r in 1:length(Dgen.sim))
for(r in 1:length(Dgen.sim))
{
  cat(r, " ")
  Results.sim[[r]] <- list()
  
  s = 3
  
  Results.sim[[r]] <- mclapply(Index.j, function(j) 
                             getMEMgene(j, Dgen = Dgen.sim[[r]][[s]], subset=NULL),
                             mc.cores=detectCores())
  names(Results.sim[[r]]) <- Sites.R.90
}

end_time <- Sys.time()
end_time - start_time
cat("This job took", (end_time - start_time)/length(Index.j)/length(loci)/R, "per dataset.")

saveRDS(Results.sim, paste0(here::here(), "/output/Results.sim.rds"))
```

## 3. Compile results for n = 90

```{r}
#Results.R.90 <- readRDS(paste0(here::here(), "/output/Results.R.90.rds"))

#Dgen.R.90 <- readRDS(paste0(here::here(), "/output/Dgen.R.90.rds"))
```

### a) Function to extract response variables

```{r}
getRes <- function(Results=Results.R.90[[1]], k = 1)
{
  RsqAdjPos = sapply(Results, function(ls) ls[[k]]$RsqAdjPos)
  RsqPos = sapply(Results, function(ls) ls[[k]]$RsqPos)
  Morans.I = sapply(Results, function(ls) ls[[k]]$Morans.I)
  I.max = sapply(Results, function(ls) max(ls[[k]]$Range))
  I.min = sapply(Results, function(ls) min(ls[[k]]$Range))
  I.scaled = Morans.I / I.max
  res <- data.frame(RsqAdjPos=RsqAdjPos, RsqPos=RsqPos, Morans.I=Morans.I, I.scaled=I.scaled)
}
```


### b) Combine response variables with Design matrix

```{r}
Results.table <- list()

for(s in length(loci):1)
{
  cat(s)
  Ftab <- data.frame(Fst.90 = sapply(Dgen.R.90[[s]], function(ls) ls$Fst[7]),
                   Fst.30.m = sapply(Dgen.R.90[[s]], function(ls) mean(ls$Fst.30[,7])),
                   Fst.30.s = sapply(Dgen.R.90[[s]], function(ls) sd(ls$Fst.30[,7])))

  res <- lapply(c(1:length(Results.R.90[[s]][[1]])), function(k) getRes(Results.R.90[[s]], k))
  names(res) <- names(Results.R.90[[s]][[1]])

  res.combined <- Reduce(cbind, res)
  names(res.combined) <- paste(rep(names(Results.R.90[[s]][[1]]), each=ncol(res[[1]])), 
                             names(res[[1]]), "90", sep=".")
  Results.table[[s]] <- data.frame(Design[Sites.R.90,], Ftab, res.combined)
}
names(Results.table) <- paste0("L", loci)

saveRDS(Results.table, paste0(here::here(), "/output/Results.table.rds"))
```

Analyze according to the following factors:

- Demography (1R, 2R, IBD, IM)
- NumInd: 20, 6
- Genetic distance measure (Fst, Dch)
- Number of loci (9900, 3300, 500, 100, 50)

Response variables:

- RsqAdjPos: as RsqAdj in memgene output with default settings (only positive MEM)
- RsqPos: unadjusted (not reported in memgene)
- Morans.I: (not reported in memgene) assuming population is sampled. Correct with (n-1)/n (where n = 90) for estimating population Moran's I
- Morans.I.scaled: this is experimental. Divide Morans.I by the maximum value (max(Range)).

Notes:

- Env: for each combination, there are three replicate datasets (Env: 453, 988, 950). Env could be used as a blocking variable.

### c) NEW Dec 2020: Standardized effect sizes

Extract simulated responses, calculate mean and sdev

```{r}
Results.table.sim <- list()

for(r in 1:length(Results.sim))
{
  Results.table.sim[[r]] <- lapply(c(1:length(Results.sim[[r]][[1]])), 
                                          function(k) getRes(Results.sim[[r]], k))
}
RsqAdjPos.sim.Fst <- data.frame(t(sapply(Results.table.sim, function(ls) ls[[1]]$RsqAdjPos)))
RsqPos.sim.Fst <- data.frame(t(sapply(Results.table.sim, function(ls) ls[[1]]$RsqPos)))
Morans.I.sim.Fst <- data.frame(t(sapply(Results.table.sim, function(ls) ls[[1]]$Morans.I)))
I.scaled.sim.Fst <- data.frame(t(sapply(Results.table.sim, function(ls) ls[[1]]$I.scaled)))

Sim.distributions <- data.frame(RsqAdjPos.mean = apply(RsqAdjPos.sim.Fst, 2, mean), 
           RsqAdjPos.sd = apply(RsqPos.sim.Fst, 2, sd),
           RsqPos.mean = apply(RsqPos.sim.Fst, 2, mean), 
           RsqPos.sd = apply(RsqAdjPos.sim.Fst, 2, sd),
           Morans.I.mean = apply(Morans.I.sim.Fst, 2, mean), 
           Morans.I.sd = apply(Morans.I.sim.Fst, 2, sd),
           I.scaled.mean = apply(I.scaled.sim.Fst, 2, mean), 
           I.scaled.sd = apply(I.scaled.sim.Fst, 2, sd))
row.names(Sim.distributions) <- names(Results.sim[[1]])

saveRDS(Sim.distributions, paste0(here::here(), "/output/Sim.distributions.rds"))
```

For each response, subtract mean of simulated values and divide by their sdev (Fst only)

```{r}
getSES <- function(Table=Results.table[[3]], Sim=Sim.distributions)
{
  Table$SES.RsqAdjPos.90 = (Table$Fst.RsqAdjPos.90 - Sim$RsqAdjPos.mean) / Sim$RsqAdjPos.sd
  Table$SES.RsqPos.90 = (Table$Fst.RsqPos.90 - Sim$RsqPos.mean) / Sim$RsqPos.sd
  Table$SES.Morans.I.90 = (Table$Fst.Morans.I.90 - Sim$Morans.I.mean) / Sim$Morans.I.sd
  Table$SES.I.scaled.90 = (Table$Fst.I.scaled.90 - Sim$I.scaled.mean) / Sim$I.scaled.sd
  Table
}
```

```{r}
#Results.table <- readRDS(paste0(here::here(), "/output/Results.table.rds"))

s = 3

Results.table[[s]] <- getSES(Table=Results.table[[s]], Sim=Sim.distributions)

saveRDS(Results.table, paste0(here::here(), "/output/Results.table.rds"))
```


## 4. Analyze data for subsets with 30 pops

### a) Repeat MEMgene with n = 30

The R = 10 subsets of n = 30 sites for each dataset with n = 90 sites were defined above already in object `Sites.30`.

```{r MEMgene30}
start_time <- Sys.time()

Index.j <- 1:length(Sites.R.90)
RR = length(Sites.30)
Results.R.30 <- list()

for(s in 1:length(Dgen.R.90))
{
  cat("\n", s, ": ")
  Results.R.30[[s]] <- rep( list(list()), RR)
  for(rr in 1:RR)
  {
    cat(rr)
    Results.R.30[[s]][[rr]] <- mclapply(Index.j, function(j) getMEMgene(j, Dgen = Dgen.R.90[[s]],
                                                         subset=Sites.30[[rr]]),
                                  mc.cores=detectCores())
    names(Results.R.30[[s]][[rr]]) <- Sites.R.90
  }
}
names(Results.R.30) <- names(Dgen.R.90)


end_time <- Sys.time()
end_time - start_time
cat("This job took", (end_time - start_time)/length(Index.j)/length(Dgen.R.90), "per dataset.")

saveRDS(Results.R.30, paste0(here::here(), "/output/Results.R.30.rds"))
```

Repeat for simulated data (only for 500 loci)

```{r}
start_time <- Sys.time()

Index.j <- 1:length(Sites.R.90)
RR = length(Sites.30)
Results.R.30.sim <- list()

for(r in 1:length(Dgen.sim))
{
  cat(r, " ")
  Results.R.30.sim[[r]] <- list()

  s = 3
  
  Results.R.30.sim[[r]] <- rep( list(list()), RR)
  for(rr in 1:RR)
  {
    Results.R.30.sim[[r]][[rr]] <- mclapply(Index.j, 
                                          function(j) getMEMgene(j, Dgen = Dgen.sim[[r]][[s]],
                                          subset=Sites.30[[rr]]),
                                          mc.cores=detectCores())
    names(Results.R.30.sim[[r]][[rr]]) <- Sites.R.90
  }
}

end_time <- Sys.time()
end_time - start_time
cat("This job took", (end_time - start_time)/length(Index.j), "per dataset.")

saveRDS(Results.R.30.sim, paste0(here::here(), "/output/Results.R.30.sim.rds"))
```

### b) Compile results for n = 30

Extract results for each replicate subsample

```{r}
#Results.R.30 <- readRDS(paste0(here::here(), "/output/Results.R.30.rds"))

res.r <- list()
Results.table.30 <- Results.table

for(s in 1:length(Results.R.30))
{
  res.r[[s]] <- lapply(Results.R.30[[s]], function(sub) 
      lapply(c(1:length(Results.R.30[[s]][[1]][[1]])), function(k) getRes(sub, k)))
  
  Results.table.30[[s]]$Fst.RsqAdjPos.30m <- apply(data.frame(t(sapply(res.r[[s]], 
                                             function(ls) ls[[1]]$RsqAdjPos))), 2, mean)
  Results.table.30[[s]]$Fst.RsqAdjPos.30s <- apply(data.frame(t(sapply(res.r[[s]], 
                                             function(ls) ls[[1]]$RsqAdjPos))), 2, sd)
  Results.table.30[[s]]$Fst.RsqPos.30m <- apply(data.frame(t(sapply(res.r[[s]], 
                                             function(ls) ls[[1]]$RsqPos))), 2, mean)
  Results.table.30[[s]]$Fst.RsqPos.30s <- apply(data.frame(t(sapply(res.r[[s]], 
                                             function(ls) ls[[1]]$RsqPos))), 2, sd)
  Results.table.30[[s]]$Fst.Morans.I.30m <- apply(data.frame(t(sapply(res.r[[s]], 
                                             function(ls) ls[[1]]$Morans.I))), 2, mean)
  Results.table.30[[s]]$Fst.Morans.I.30s <- apply(data.frame(t(sapply(res.r[[s]], 
                                             function(ls) ls[[1]]$Morans.I))), 2, sd)
  Results.table.30[[s]]$Fst.I.scaled.30m <- apply(data.frame(t(sapply(res.r[[s]], 
                                             function(ls) ls[[1]]$I.scaled))), 2, mean)
  Results.table.30[[s]]$Fst.I.scaled.30s <- apply(data.frame(t(sapply(res.r[[s]], 
                                             function(ls) ls[[1]]$I.scaled))), 2, sd)
  Results.table.30[[s]]$Dch.RsqAdjPos.30m <- apply(data.frame(t(sapply(res.r[[s]], 
                                             function(ls) ls[[2]]$RsqAdjPos))), 2, mean)
  Results.table.30[[s]]$Dch.RsqAdjPos.30s <- apply(data.frame(t(sapply(res.r[[s]], 
                                             function(ls) ls[[2]]$RsqAdjPos))), 2, sd)
  Results.table.30[[s]]$Dch.RsqPos.30m <- apply(data.frame(t(sapply(res.r[[s]], 
                                             function(ls) ls[[2]]$RsqPos))), 2, mean)
  Results.table.30[[s]]$Dch.RsqPos.30s <- apply(data.frame(t(sapply(res.r[[s]], 
                                             function(ls) ls[[2]]$RsqPos))), 2, sd)
  Results.table.30[[s]]$Dch.Morans.I.30m <- apply(data.frame(t(sapply(res.r[[s]], 
                                             function(ls) ls[[2]]$Morans.I))), 2, mean)
  Results.table.30[[s]]$Dch.Morans.I.30s <- apply(data.frame(t(sapply(res.r[[s]], 
                                             function(ls) ls[[2]]$Morans.I))), 2, sd)
  Results.table.30[[s]]$Dch.I.scaled.30m <- apply(data.frame(t(sapply(res.r[[s]], 
                                             function(ls) ls[[2]]$I.scaled))), 2, mean)
  Results.table.30[[s]]$Dch.I.scaled.30s <- apply(data.frame(t(sapply(res.r[[s]], 
                                             function(ls) ls[[2]]$I.scaled))), 2, sd)
}
```

Repeat for simulated data, get means and sdev (Fst only)

```{r}
Results.table.sim.30 <- list()

  for(r in 1:length(Results.R.30.sim))
  {
    Results.table.sim.30[[r]] <- lapply(Results.R.30.sim[[r]], function(sub) getRes(sub, 1))
  }
  
  RsqAdjPos.sim.30.Fst <- lapply(Results.table.sim.30, function(ls) sapply(ls, 
                                 function(x) x$RsqAdjPos))
  RsqPos.sim.30.Fst <- lapply(Results.table.sim.30, function(ls) sapply(ls, 
                                 function(x) x$RsqPos))
  Morans.I.sim.30.Fst <- lapply(Results.table.sim.30, function(ls) sapply(ls, 
                                 function(x) x$Morans.I))
  I.scaled.sim.30.Fst <- lapply(Results.table.sim.30, function(ls) sapply(ls, 
                                 function(x) x$I.scaled))

  
  RsqAdjPos.30.mean <- Reduce("+", RsqAdjPos.sim.30.Fst)/length(RsqAdjPos.sim.30.Fst)
  tmp <- lapply(RsqAdjPos.sim.30.Fst, function(ls) (ls - RsqAdjPos.30.mean)^2)
  RsqAdjPos.30.sd <- sqrt(Reduce("+", tmp)/(length(tmp) - 1))
  
  RsqPos.30.mean <- Reduce("+", RsqPos.sim.30.Fst)/length(RsqPos.sim.30.Fst)
  tmp <- lapply(RsqPos.sim.30.Fst, function(ls) (ls - RsqPos.30.mean)^2)
  RsqPos.30.sd <- sqrt(Reduce("+", tmp)/(length(tmp) - 1))
  
  Morans.I.30.mean <- Reduce("+", Morans.I.sim.30.Fst)/length(Morans.I.sim.30.Fst)
  tmp <- lapply(Morans.I.sim.30.Fst, function(ls) (ls - Morans.I.30.mean)^2)
  Morans.I.30.sd <- sqrt(Reduce("+", tmp)/(length(tmp) - 1))
  
  I.scaled.30.mean <- Reduce("+", I.scaled.sim.30.Fst)/length(I.scaled.sim.30.Fst)
  tmp <- lapply(I.scaled.sim.30.Fst, function(ls) (ls - I.scaled.30.mean)^2)
  I.scaled.30.sd <- sqrt(Reduce("+", tmp)/(length(tmp) - 1))
  
  Sim.distributions.30 <- list(RsqAdjPos.30.mean = RsqAdjPos.30.mean, 
           RsqAdjPos.30.sd = RsqAdjPos.30.sd,
           RsqPos.30.mean = RsqPos.30.mean, 
           RsqPos.30.sd = RsqPos.30.sd,
           Morans.I.30.mean = Morans.I.30.mean, 
           Morans.I.30.sd = Morans.I.30.sd,
           I.scaled.30.mean = I.scaled.30.mean, 
           I.scaled.30.sd = I.scaled.30.sd)
  
saveRDS(Sim.distributions.30, paste0(here::here(), "/output/Sim.distributions.30.rds"))
```

Calculate SES for replicate subsamples

```{r}
Results.subsets.table <- Results.table.30

#for(s in length(Results.table.30):1)
s = 3

  Results.subsets.table[[s]]$SES.RsqAdjPos.30m <- 
    apply((t(sapply(res.r[[s]], function(ls) ls[[1]]$RsqAdjPos)) -
       t(Sim.distributions.30$RsqAdjPos.30.mean)) / 
       t(Sim.distributions.30$RsqAdjPos.30.sd), 2, mean)
  Results.subsets.table[[s]]$SES.RsqAdjPos.30s <- 
    apply((t(sapply(res.r[[s]], function(ls) ls[[1]]$RsqAdjPos)) -
       t(Sim.distributions.30$RsqAdjPos.30.mean)) / 
       t(Sim.distributions.30$RsqAdjPos.30.sd), 2, sd)
  Results.subsets.table[[s]]$SES.RsqPos.30m <- 
    apply((t(sapply(res.r[[s]], function(ls) ls[[1]]$RsqPos)) -
       t(Sim.distributions.30$RsqPos.30.mean)) / 
       t(Sim.distributions.30$RsqPos.30.sd), 2, mean)
  Results.subsets.table[[s]]$SES.RsqPos.30s <- 
    apply((t(sapply(res.r[[s]], function(ls) ls[[1]]$RsqPos)) -
       t(Sim.distributions.30$RsqPos.30.mean)) / 
       t(Sim.distributions.30$RsqPos.30.sd), 2, sd)
  Results.subsets.table[[s]]$SES.Morans.I.30m <- 
    apply((t(sapply(res.r[[s]], function(ls) ls[[1]]$Morans.I)) -
       t(Sim.distributions.30$Morans.I.30.mean)) / 
       t(Sim.distributions.30$Morans.I.30.sd), 2, mean)
  Results.subsets.table[[s]]$SES.Morans.I.30s <- 
    apply((t(sapply(res.r[[s]], function(ls) ls[[1]]$Morans.I)) -
       t(Sim.distributions.30$Morans.I.30.mean)) / 
       t(Sim.distributions.30$Morans.I.30.sd), 2, sd)
  Results.subsets.table[[s]]$SES.I.scaled.30m <- 
    apply((t(sapply(res.r[[s]], function(ls) ls[[1]]$I.scaled)) -
       t(Sim.distributions.30$I.scaled.30.mean)) / 
       t(Sim.distributions.30$I.scaled.30.sd), 2, mean)
  Results.subsets.table[[s]]$SES.I.scaled.30s <- 
    apply((t(sapply(res.r[[s]], function(ls) ls[[1]]$I.scaled)) -
       t(Sim.distributions.30$I.scaled.30.mean)) / 
       t(Sim.distributions.30$I.scaled.30.sd), 2, sd)


saveRDS(Results.subsets.table, paste0(here::here(), "/output/Results.subsets.table.rds"))  
```


## 5. Compare n = 30 to n = 90

### c) Paired t-tests  (without IM)

Drop IM:

```{r}
a <- which(Results.subsets.table[[3]]$Demography != "IM")
```


Paired t-tests for difference between mean of subsamples with n = 30 and corresponding datasets with n = 90 sites.

```{r}
s = 1

with(Results.subsets.table[[s]][a,], t.test(Fst.RsqAdjPos.30m, Fst.RsqAdjPos.90, paired=TRUE))

with(Results.subsets.table[[s]][a,], t.test(Fst.Morans.I.30m, Fst.Morans.I.90, paired=TRUE))

with(Results.subsets.table[[s]][a,], t.test(Fst.I.scaled.30m, Fst.I.scaled.90, paired=TRUE))
```

Repeat for SES (with 500 loci)

```{r}
s = 3  # For SES

with(Results.subsets.table[[s]][a,], t.test(SES.RsqAdjPos.30m, SES.RsqAdjPos.90, paired=TRUE))

with(Results.subsets.table[[s]][a,], t.test(SES.Morans.I.30m, SES.RsqAdjPos.90, paired=TRUE))

with(Results.subsets.table[[s]][a,], t.test(SES.I.scaled.30m, SES.RsqAdjPos.90, paired=TRUE))
```


### d) Effect sizes (without IM)

Cohen's d for paired t-test. 

```{r}
s=1

with(Results.subsets.table[[s]][a,], mean(Fst.RsqAdjPos.30m - Fst.RsqAdjPos.90)/ 
       sd(Fst.RsqAdjPos.30m - Fst.RsqAdjPos.90))

with(Results.subsets.table[[s]][a,], mean(Fst.Morans.I.30m - Fst.Morans.I.90)/
       sd(Fst.Morans.I.30m - Fst.Morans.I.90)) 

with(Results.subsets.table[[s]][a,], mean(Fst.I.scaled.30m - Fst.I.scaled.90)/
       sd(Fst.I.scaled.30m - Fst.I.scaled.90))
```


## 6. Missing sites (no permutations)

Simulations to compare results between full sample and sample with some missing sites. Two possible methods:

- Meta-analysis mode: redo MEMgene (including MEM) for each dataset, based on available sites
- Comparative mode: keep MEM, do adj R2 based on positive MEM, which avoids overfitting

### a) Leave 30 sites out at a time, two modes

Comparative mode: use MEM from full dataset
Meta-analysis mode: MEM based on subset of 60 sites

For 500 loci:

```{r MEMgeneDrop}
start_time <- Sys.time()

Index.j <- 1:length(Sites.R.90)
RR = 10
Results.drop <- Results.drop.MEM <- rep( list(list()), R)

set.seed(297)
Drop <-lapply(c(1:R), function(r) sort(sample(1:90, 30)))

for(rr in 1:RR)
{
  cat(rr)
  Results.drop[[rr]] <- mclapply(Index.j, function(j) getMEMgene(j, Dgen = Dgen.R.90[[3]],
                                                       subset=NULL, drop=Drop[[rr]]),
                                mc.cores=detectCores())
   Results.drop.MEM[[rr]] <- mclapply(Index.j, function(j) getMEMgene(j, Dgen = Dgen.R.90[[3]],
                                                       subset=c(1:90)[-Drop[[rr]]], drop=NULL),
                                mc.cores=detectCores())
  names(Results.drop[[rr]]) <- names(Results.drop.MEM[[rr]]) <- Sites.R.90
}

end_time <- Sys.time()
end_time - start_time
cat("This job took", (end_time - start_time)/length(Index.j), "per dataset.")

saveRDS(Results.drop, paste0(here::here(), "/output/Results.drop.rds"))
saveRDS(Results.drop.MEM, paste0(here::here(), "/output/Results.drop.MEM.rds"))
```

Repeat for simulated data: comparative and meta-analysis modes (takes 1 min per simulation run)

```{r}
start_time <- Sys.time()

Index.j <- 1:length(Sites.R.90)
RR = 10

Results.drop.sim <- Results.drop.sim.MEM <- list()

for(r in 1:length(Dgen.sim))
{
  cat("\n", r, " ")
  Results.drop.sim[[r]] <- Results.drop.sim.MEM[[r]] <- list()
  for(rr in 1:RR)
  {
    cat(rr)
    Results.drop.sim[[r]][[rr]] <- mclapply(Index.j, function(j) 
                                    getMEMgene(j, Dgen = Dgen.sim[[r]][[3]],
                                               subset=NULL, drop=Drop[[rr]]),
                                  mc.cores=detectCores())
    Results.drop.sim.MEM[[r]][[rr]] <- mclapply(Index.j, function(j) 
                                    getMEMgene(j, Dgen = Dgen.sim[[r]][[3]],
                                               subset=c(1:90)[-Drop[[rr]]], drop=NULL),
                                  mc.cores=detectCores())
    names(Results.drop.sim[[r]][[rr]]) <-  names(Results.drop.sim.MEM[[r]][[rr]]) <- Sites.R.90
  }
}

end_time <- Sys.time()
end_time - start_time
cat("This job took", (end_time - start_time)/length(Index.j), "per dataset.")

saveRDS(Results.drop.sim, paste0(here::here(), "/output/Results.drop.sim.rds"))
saveRDS(Results.drop.sim.MEM, paste0(here::here(), "/output/Results.drop.MEM.rds"))
```


Get means and sdev for simulated data(500 loci, Fst only)

```{r}
#Results.drop.sim <- readRDS(paste0(here::here(), "/output/Results.drop.sim.rds"))
#Results.drop.sim.MEM <- readRDS(paste0(here::here(), "/output/Results.drop.MEM.rds"))

Results.table.drop.sim <- list()
Results.table.drop.sim.MEM <- list()

#for(s in length(loci):1)
s = 3
  for(r in 1:length(Results.drop.sim))
  {
    Results.table.drop.sim[[r]] <- lapply(Results.drop.sim[[r]], function(sub)
                    getRes(sub, 1))
    Results.table.drop.sim.MEM[[r]] <- lapply(Results.drop.sim.MEM[[r]], function(sub)
                    getRes(sub, 1))
  }
  
  RsqAdjPos.drop.sim.Fst <- lapply(Results.table.drop.sim, function(ls) 
                                 sapply(ls, function(x) x$RsqAdjPos))
  RsqPos.drop.sim.Fst <- lapply(Results.table.drop.sim, function(ls) 
                                 sapply(ls, function(x) x$RsqPos))
  Morans.I.drop.sim.Fst <- lapply(Results.table.drop.sim, function(ls) 
                                 sapply(ls, function(x) x$Morans.I))
  I.scaled.drop.sim.Fst <- lapply(Results.table.drop.sim, function(ls) 
                                 sapply(ls, function(x) x$I.scaled))
  RsqAdjPos.drop.sim.MEM.Fst <- lapply(Results.table.drop.sim.MEM, function(ls) 
                                 sapply(ls, function(x) x$RsqAdjPos))
  RsqPos.drop.sim.MEM.Fst <- lapply(Results.table.drop.sim.MEM, function(ls) 
                                 sapply(ls, function(x) x$RsqPos))
  Morans.I.drop.sim.MEM.Fst <- lapply(Results.table.drop.sim.MEM, function(ls) 
                                 sapply(ls, function(x) x$Morans.I))
  I.scaled.drop.sim.MEM.Fst <- lapply(Results.table.drop.sim.MEM, function(ls) 
                                 sapply(ls, function(x) x$I.scaled))

  RsqAdjPos.drop.sim.mean <- Reduce("+", RsqAdjPos.drop.sim.Fst)/length(RsqAdjPos.drop.sim.Fst)
  tmp <- lapply(RsqAdjPos.drop.sim.Fst, function(ls) (ls - RsqAdjPos.drop.sim.mean)^2)
  RsqAdjPos.drop.sim.sd <- sqrt(Reduce("+", tmp)/(length(tmp) - 1))
  
  RsqPos.drop.sim.mean <- Reduce("+", RsqPos.drop.sim.Fst)/length(RsqPos.drop.sim.Fst)
  tmp <- lapply(RsqPos.drop.sim.Fst, function(ls) (ls - RsqPos.drop.sim.mean)^2)
  RsqPos.drop.sim.sd <- sqrt(Reduce("+", tmp)/(length(tmp) - 1))
  
  Morans.I.drop.sim.mean <- Reduce("+", Morans.I.drop.sim.Fst)/length(Morans.I.drop.sim.Fst)
  tmp <- lapply(Morans.I.drop.sim.Fst, function(ls) (ls - Morans.I.drop.sim.mean)^2)
  Morans.I.drop.sim.sd <- sqrt(Reduce("+", tmp)/(length(tmp) - 1))
  
  I.scaled.drop.sim.mean <- Reduce("+", I.scaled.drop.sim.Fst)/length(I.scaled.drop.sim.Fst)
  tmp <- lapply(I.scaled.drop.sim.Fst, function(ls) (ls - I.scaled.drop.sim.mean)^2)
  I.scaled.drop.sim.sd <- sqrt(Reduce("+", tmp)/(length(tmp) - 1))
  
  RsqAdjPos.drop.sim.MEM.mean <- Reduce("+", RsqAdjPos.drop.sim.MEM.Fst)/
                                             length(RsqAdjPos.drop.sim.MEM.Fst)
  tmp <- lapply(RsqAdjPos.drop.sim.MEM.Fst, function(ls) (ls - RsqAdjPos.drop.sim.MEM.mean)^2)
  RsqAdjPos.drop.sim.MEM.sd <- sqrt(Reduce("+", tmp)/(length(tmp) - 1))
  
  RsqPos.drop.sim.MEM.mean <- Reduce("+", RsqPos.drop.sim.MEM.Fst)/
                                             length(RsqPos.drop.sim.MEM.Fst)
  tmp <- lapply(RsqPos.drop.sim.MEM.Fst, function(ls) (ls - RsqPos.drop.sim.MEM.mean)^2)
  RsqPos.drop.sim.MEM.sd <- sqrt(Reduce("+", tmp)/(length(tmp) - 1))
  
  Morans.I.drop.sim.MEM.mean <- Reduce("+", Morans.I.drop.sim.MEM.Fst)/
                                            length(Morans.I.drop.sim.MEM.Fst)
  tmp <- lapply(Morans.I.drop.sim.MEM.Fst, function(ls) (ls - Morans.I.drop.sim.MEM.mean)^2)
  Morans.I.drop.sim.MEM.sd <- sqrt(Reduce("+", tmp)/(length(tmp) - 1))
  
  I.scaled.drop.sim.MEM.mean <- Reduce("+", I.scaled.drop.sim.MEM.Fst)/
                                            length(I.scaled.drop.sim.MEM.Fst)
  tmp <- lapply(I.scaled.drop.sim.MEM.Fst, function(ls) (ls - I.scaled.drop.sim.MEM.mean)^2)
  I.scaled.drop.sim.MEM.sd <- sqrt(Reduce("+", tmp)/(length(tmp) - 1))
  
  Sim.drop.distributions <- list(RsqAdjPos.drop.sim.mean = RsqAdjPos.drop.sim.mean, 
           RsqAdjPos.drop.sim.sd = RsqAdjPos.drop.sim.sd,
           RsqPos.drop.sim.mean = RsqPos.drop.sim.mean, 
           RsqPos.drop.sim.sd = RsqPos.drop.sim.sd,
           Morans.I.drop.sim.mean = Morans.I.drop.sim.mean, 
           Morans.I.drop.sim.sd = Morans.I.drop.sim.sd,
           I.scaled.drop.sim.mean = I.scaled.drop.sim.mean, 
           I.scaled.drop.sim.sd = I.scaled.drop.sim.sd,
           
           RsqAdjPos.drop.sim.MEM.mean = RsqAdjPos.drop.sim.MEM.mean, 
           RsqAdjPos.drop.sim.MEM.sd = RsqAdjPos.drop.sim.MEM.sd,
           RsqPos.drop.sim.MEM.mean = RsqPos.drop.sim.MEM.mean, 
           RsqPos.drop.sim.MEM.sd = RsqPos.drop.sim.MEM.sd,
           Morans.I.drop.sim.MEM.mean = Morans.I.drop.sim.MEM.mean, 
           Morans.I.drop.sim.MEM.sd = Morans.I.drop.sim.MEM.sd,
           I.scaled.drop.sim.MEM.mean = I.scaled.drop.sim.MEM.mean, 
           I.scaled.drop.sim.MEM.sd = I.scaled.drop.sim.MEM.sd)

  saveRDS(Sim.drop.distributions, paste0(here::here(), "/output/Sim.drop.distributions.rds"))
```

### b) Compile results for n = 60, two modes

Determine mean and sdev among the R = 10 replicate subsets for each dataset with 5 sites dropped


```{r}
#Results.drop <- readRDS(paste0(here::here(), "/output/Results.drop.rds"))
#Results.drop.MEM <- readRDS(paste0(here::here(), "/output/Results.drop.MEM.rds"))
#Results.subsets.table <- readRDS(paste0(here::here(), "/output/Results.subsets.table.rds"))

res.r <- lapply(Results.drop, function(sub) 
  lapply(c(1:length(Results.drop[[1]][[1]])), function(k) getRes(sub, k)))
res.r.MEM <- lapply(Results.drop.MEM, function(sub) 
  lapply(c(1:length(Results.drop.MEM[[1]][[1]])), function(k) getRes(sub, k)))
```


```{r}

tmp <- list()
for(k in 1:length(res.r[[1]]))
  {
    RsqAdjPos.drop.m <- apply(sapply(res.r, function(sub) sub[[k]]$RsqAdjPos), 1, mean)
    RsqAdjPos.drop.s <- apply(sapply(res.r, function(sub) sub[[k]]$RsqAdjPos), 1, sd)
    RsqPos.drop.m <- apply(sapply(res.r, function(sub) sub[[k]]$RsqPos), 1, mean)
    RsqPos.drop.s <- apply(sapply(res.r, function(sub) sub[[k]]$RsqPos), 1, sd)
    Morans.I.drop.m <- apply(sapply(res.r, function(sub) sub[[k]]$Morans.I), 1, mean)
    Morans.I.drop.s <- apply(sapply(res.r, function(sub) sub[[k]]$Morans.I), 1, sd)
    I.scaled.drop.m <- apply(sapply(res.r, function(sub) sub[[k]]$I.scaled), 1, mean)
    I.scaled.drop.s <- apply(sapply(res.r, function(sub) sub[[k]]$I.scaled), 1, sd)
    
    RsqAdjPos.drop.MEM.m <- apply(sapply(res.r.MEM, function(sub) sub[[k]]$RsqAdjPos), 1, mean)
    RsqAdjPos.drop.MEM.s <- apply(sapply(res.r.MEM, function(sub) sub[[k]]$RsqAdjPos), 1, sd)
    RsqPos.drop.MEM.m <- apply(sapply(res.r.MEM, function(sub) sub[[k]]$RsqPos), 1, mean)
    RsqPos.drop.MEM.s <- apply(sapply(res.r.MEM, function(sub) sub[[k]]$RsqPos), 1, sd)
    Morans.I.drop.MEM.m <- apply(sapply(res.r.MEM, function(sub) sub[[k]]$Morans.I), 1, mean)
    Morans.I.drop.MEM.s <- apply(sapply(res.r.MEM, function(sub) sub[[k]]$Morans.I), 1, sd)
    I.scaled.drop.MEM.m <- apply(sapply(res.r.MEM, function(sub) sub[[k]]$I.scaled), 1, mean)
    I.scaled.drop.MEM.s <- apply(sapply(res.r.MEM, function(sub) sub[[k]]$I.scaled), 1, sd)
  
    tmp[[k]] <- data.frame(RsqAdjPos.drop.m=RsqAdjPos.drop.m, RsqAdjPos.drop.s=RsqAdjPos.drop.s, 
                           RsqPos.drop.m=RsqPos.drop.m, RsqPos.drop.s=RsqPos.drop.s,
                           Morans.I.drop.m=Morans.I.drop.m, Morans.I.drop.s=Morans.I.drop.s,
                           I.scaled.drop.m=I.scaled.drop.m, I.scaled.drop.s=I.scaled.drop.s,
                           
                           RsqAdjPos.drop.MEM.m=RsqAdjPos.drop.MEM.m, 
                           RsqAdjPos.drop.MEM.s=RsqAdjPos.drop.MEM.s, 
                           RsqPos.drop.MEM.m=RsqPos.drop.MEM.m, 
                           RsqPos.drop.MEM.s=RsqPos.drop.MEM.s,
                           Morans.I.drop.MEM.m=Morans.I.drop.MEM.m,
                           Morans.I.drop.MEM.s=Morans.I.drop.MEM.s,
                           I.scaled.drop.MEM.m=I.scaled.drop.MEM.m, 
                           I.scaled.drop.MEM.s=I.scaled.drop.MEM.s)
}

res.r.combined <- Reduce(cbind, tmp)

names(res.r.combined) <- paste(rep(names(Results.drop[[1]][[1]]), each=ncol(tmp[[1]])), 
                             names(tmp[[1]]), sep=".")

Results.subsets.table[[3]] <- data.frame(Results.subsets.table[[3]], res.r.combined)
```

Calculate SES

```{r}
s = 3

  Results.subsets.table[[s]]$SES.RsqAdjPos.drop.m <- 
    apply((t(sapply(res.r, function(ls) ls[[1]]$RsqAdjPos)) -
       t(Sim.drop.distributions$RsqAdjPos.drop.sim.mean)) / 
       t(Sim.drop.distributions$RsqAdjPos.drop.sim.sd), 2, mean)
  Results.subsets.table[[s]]$SES.RsqAdjPos.drop.s <- 
    apply((t(sapply(res.r, function(ls) ls[[1]]$RsqAdjPos)) -
       t(Sim.drop.distributions$RsqAdjPos.drop.sim.mean)) / 
       t(Sim.drop.distributions$RsqAdjPos.drop.sim.sd), 2, sd)
  
  Results.subsets.table[[s]]$SES.RsqPos.drop.m <- 
    apply((t(sapply(res.r, function(ls) ls[[1]]$RsqPos)) -
       t(Sim.drop.distributions$RsqPos.drop.sim.mean)) / 
       t(Sim.drop.distributions$RsqPos.drop.sim.sd), 2, mean)
  Results.subsets.table[[s]]$SES.RsqPos.drop.s <- 
    apply((t(sapply(res.r, function(ls) ls[[1]]$RsqPos)) -
       t(Sim.drop.distributions$RsqPos.drop.sim.mean)) / 
       t(Sim.drop.distributions$RsqPos.drop.sim.sd), 2, sd)
  
  Results.subsets.table[[s]]$SES.Morans.I.drop.m <- 
    apply((t(sapply(res.r, function(ls) ls[[1]]$Morans.I)) -
       t(Sim.drop.distributions$Morans.I.drop.sim.mean)) / 
       t(Sim.drop.distributions$Morans.I.drop.sim.sd), 2, mean)
  Results.subsets.table[[s]]$SES.Morans.I.drop.s <- 
    apply((t(sapply(res.r, function(ls) ls[[1]]$Morans.I)) -
       t(Sim.drop.distributions$Morans.I.drop.sim.mean)) / 
       t(Sim.drop.distributions$Morans.I.drop.sim.sd), 2, sd)
  
  Results.subsets.table[[s]]$SES.I.scaled.drop.m <- 
    apply((t(sapply(res.r, function(ls) ls[[1]]$I.scaled)) -
       t(Sim.drop.distributions$I.scaled.drop.sim.mean)) / 
       t(Sim.drop.distributions$I.scaled.drop.sim.sd), 2, mean)
  Results.subsets.table[[s]]$SES.I.scaled.drop.s <- 
    apply((t(sapply(res.r, function(ls) ls[[1]]$I.scaled)) -
       t(Sim.drop.distributions$I.scaled.drop.sim.mean)) / 
       t(Sim.drop.distributions$I.scaled.drop.sim.sd), 2, sd)
  
  Results.subsets.table[[s]]$SES.RsqAdjPos.drop.MEM.m <- 
    apply((t(sapply(res.r, function(ls) ls[[1]]$RsqAdjPos)) -
       t(Sim.drop.distributions$RsqAdjPos.drop.sim.MEM.mean)) / 
       t(Sim.drop.distributions$RsqAdjPos.drop.sim.MEM.sd), 2, mean)
  Results.subsets.table[[s]]$SES.RsqAdjPos.drop.MEM.s <- 
    apply((t(sapply(res.r, function(ls) ls[[1]]$RsqAdjPos)) -
       t(Sim.drop.distributions$RsqAdjPos.drop.sim.MEM.mean)) / 
       t(Sim.drop.distributions$RsqAdjPos.drop.sim.MEM.sd), 2, sd)
  
  Results.subsets.table[[s]]$SES.RsqPos.drop.MEM.m <- 
    apply((t(sapply(res.r, function(ls) ls[[1]]$RsqPos)) -
       t(Sim.drop.distributions$RsqPos.drop.sim.MEM.mean)) / 
       t(Sim.drop.distributions$RsqPos.drop.sim.MEM.sd), 2, mean)
  Results.subsets.table[[s]]$SES.RsqPos.drop.MEM.s <- 
    apply((t(sapply(res.r, function(ls) ls[[1]]$RsqPos)) -
       t(Sim.drop.distributions$RsqPos.drop.sim.MEM.mean)) / 
       t(Sim.drop.distributions$RsqPos.drop.sim.MEM.sd), 2, sd)
  
  Results.subsets.table[[s]]$SES.Morans.I.drop.MEM.m <- 
    apply((t(sapply(res.r, function(ls) ls[[1]]$Morans.I)) -
       t(Sim.drop.distributions$Morans.I.drop.sim.MEM.mean)) / 
       t(Sim.drop.distributions$Morans.I.drop.sim.MEM.sd), 2, mean)
  Results.subsets.table[[s]]$SES.Morans.I.drop.MEM.s <- 
    apply((t(sapply(res.r, function(ls) ls[[1]]$Morans.I)) -
       t(Sim.drop.distributions$Morans.I.drop.sim.MEM.mean)) / 
       t(Sim.drop.distributions$Morans.I.drop.sim.MEM.sd), 2, sd)
  
  Results.subsets.table[[s]]$SES.I.scaled.drop.MEM.m <- 
    apply((t(sapply(res.r, function(ls) ls[[1]]$I.scaled)) -
       t(Sim.drop.distributions$I.scaled.drop.sim.MEM.mean)) / 
       t(Sim.drop.distributions$I.scaled.drop.sim.MEM.sd), 2, mean)
  Results.subsets.table[[s]]$SES.I.scaled.drop.MEM.s <- 
    apply((t(sapply(res.r, function(ls) ls[[1]]$I.scaled)) -
       t(Sim.drop.distributions$I.scaled.drop.sim.MEM.mean)) / 
       t(Sim.drop.distributions$I.scaled.drop.sim.MEM.sd), 2, sd)


saveRDS(Results.subsets.table, paste0(here::here(), "/output/Results.subsets.table.rds"))  
```


## 7. Summary Interaction Plots

```{r}
library("rcompanion")
library("ggpubr")
library("ggplot2")
library("gridExtra")
library("reshape2")
library("tidyr")

source(paste0(here::here(), "/output/InteractionPlotLegend.R"))
```

Import results table

```{r}
#Results.table <- readRDS(paste0(here::here(), "/output/Results.drop.MEM.table.rds"))
```

### a) Figure 1

With n=90, #IND= 20 & the 9900 loci: Difference demographies in response variabe (Fst, adjusted R^2 and Moran's I).

```{r fig.height=2.5, fig.width=6}

#Results.table <- readRDS(paste0(here::here(), "/output/Results.table.rds"))
#Results.subsets.table <- readRDS(paste0(here::here(), "/output/Results.subsets.table.rds"))
#Results.table <- Results.subsets.table

s = 1
b <- which(Results.table[[s]]$NumInd == 20)
Results.table[[s]]$Demography <- factor(Results.table[[s]]$Demography, 
                                        levels = c("IM", "IBD", "1R", "2R"))

Sum1.p <- groupwiseMean(Fst.90 ~ Demography, data = Results.table[[s]][b,], 
                        conf   = 0.95, digits = 3)
Pan.a.p<- ggplot(Sum1.p, aes(x = Demography, y = Mean)) + ylim(0.0,0.02) +
   geom_errorbar(aes(ymin = Trad.lower, ymax = Trad.upper), width = 0.05, size  = 0.5) +
   geom_point(shape = 1, size  = 2) +
   theme_bw() + theme (panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
    ggtitle("(a)") + xlab("Demography") + ylab("Fst") +
    theme(axis.text = element_text(size = 12),
          plot.title = element_text(size=16, face="bold"),
          axis.title.x = element_text(size=12, face="bold"),
          axis.title.y = element_text(size=12, face="bold")
          )
#Pan.a.p

Sum2.p <- groupwiseMean(Fst.RsqAdjPos.90 ~ Demography, data = Results.table[[s]][b,], 
                        conf   = 0.95, digits = 3)
Pan.b.p<- ggplot(Sum2.p, aes(x = Demography, y = Mean)) + ylim(-0.02,0.8) +
   geom_errorbar(aes(ymin = Trad.lower, ymax = Trad.upper), width = 0.05, size  = 0.5) +
   geom_point(shape = 1, size  = 2) +
   theme_bw() + theme (panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
    ggtitle("(b)") + xlab("Demography") + ylab(expression(bold('Adjusted R'^2))) +
    theme(axis.text = element_text(size = 12),
          plot.title = element_text(size=16, face="bold"),
          axis.title.x = element_text(size=12, face="bold"),
          axis.title.y = element_text(size=12, face="bold")
          )

Sum3.p <- groupwiseMean(Fst.Morans.I.90 ~ Demography, data = Results.table[[s]][b,], 
                        conf   = 0.95, digits = 3)
Pan.c.p<- ggplot(Sum3.p, aes(x = Demography, y = Mean)) + ylim(-0.02,0.8) +
   geom_errorbar(aes(ymin = Trad.lower, ymax = Trad.upper), width = 0.05, size  = 0.5) +
   geom_point(shape = 1, size  = 2) +
   theme_bw() + theme (panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
    ggtitle("(c)") + xlab("Demography") + ylab("Moran's I") +
    theme(axis.text = element_text(size = 12),
          plot.title = element_text(size=16, face="bold"),
          axis.title.x = element_text(size=12, face="bold"),
          axis.title.y = element_text(size=12, face="bold")
          )
figure.1.P <- ggarrange(Pan.a.p, Pan.b.p, Pan.c.p, ncol = 3, nrow = 1)
#figure.1.P

g <- arrangeGrob(Pan.a.p, Pan.b.p, Pan.c.p, ncol = 3, nrow = 1) #generates g
 ggsave(file=paste0(here::here(), "/output/Figure1.pdf"), g)
 ggsave(file=paste0(here::here(), "/output/Figure1.jpg"), g)

```


### b) Figure 2

With n=90 and removing IM: Interaction between #Ind (20, 6) and # Loci (9900, 3300, 500) in response variabe (Fst, adjusted R^2, and Moran's I).


```{r fig.height=2.5, fig.width=6}
Results.table <- Results.subsets.table 

a <- which(Results.table[[1]]$Demography != "IM")

tmp <- lapply(1:length(Results.table), function(s) 
  data.frame(NumLoci=rep(loci[s],length(a)), Results.table[[s]] %>% 
             mutate(Demography = as.character(Demography), 
                    NumInd = as.character(NumInd)) %>%
             filter(Demography!= "IM") %>%   
             select(Demography, NumPops, NumInd, Fst.90, Fst.RsqAdjPos.90, 
                    Fst.Morans.I.90, Fst.I.scaled.90)))

Results.sub.long <- Reduce(rbind, tmp)
Results.sub.long <- Results.sub.long %>% 
            mutate(NumLoci = factor(NumLoci, levels = c("50", "100", "500", "3300", "9900"))) %>%
            mutate(NumInd = factor(NumInd, levels = c("20", "6")))
names(Results.sub.long)[5:8] <- c("Fst", "Adj.Rsqr.Pos", "Morans.I", "I.scaled")


#pdf(file = paste0(here::here(), "/output/Figure2.pdf"), width=8, height=3.5)
jpeg(file = paste0(here::here(), "/output/Figure2.jpg"), width=8, height=3.5, units="in", res=300)

par(mfcol=c(1,3), mai = c(0.8, 0.8, 0.5, 0.1), omi=c(0,0,0,0))
  with(Results.sub.long, {interaction.plot(NumInd, NumLoci, Fst, 
                            col=c("grey70","grey50","grey30", "grey20", "grey1"), 
                            ylab=" Mean Fst", ylim=c(0.0,0.015), lwd= 1.5, 
                            xlab="Number of individuals", cex.lab=1.5, cex.axis=1.5, 
                            cex.sub=1.5, xleg = "bottomright")})
  title("(a)", adj = 0, line =1, cex.main=2 )
    
 with(Results.sub.long, {interaction.plot(NumInd, NumLoci, Adj.Rsqr.Pos, 
                           col=c("grey70","grey50","grey30", "grey20", "grey1"), 
                           ylab=expression("Mean adjusted R"^2), ylim=c(0.0,0.6), 
                           xlab="Number of individuals", lwd= 1.5, cex.lab=1.5, 
                           cex.axis=1.5, cex.sub=1.5, legend=FALSE)})
 title("(b)", adj = 0, line =1, cex.main=2 )
  
 with(Results.sub.long, {interaction.plot(NumInd, NumLoci, Morans.I, 
                           col=c("grey70","grey50","grey30", "grey20", "grey1"), 
                           ylab="Mean Moran's I", ylim=c(0.0,0.6), 
                           xlab="Number of individuals", lwd= 1.5, cex.lab=1.5, 
                           cex.axis=1.5, cex.sub=1.5, legend=FALSE)})
 title("(c)", adj = 0, line =1, cex.main=2 )

dev.off()

```

### c) Figure 3


Multipanel fig. one figure for each demography all on the same scale. boxplots (20 ind & 500 loci)

Respo= Adj R^2 & MI & scaled MI. (x axis = 60 refitting, missing values, 30 refitting)... show variability as proportion (divide each value by the value of the 90)

NEW DEC 2020: 

- Using 500 instead of 9900 loci. 
- Added SES.
- Added unadjusted Rsquared, as SES for adjusted Rsquared problematic.

```{r}
Results.table <- Results.subsets.table[[3]] %>% filter(NumInd == 20)
Results.table <- split(Results.table, Results.table$Demography)

#Plots.data <- list()
Plots <- list()

demo <- rep(c("IBD", "1R", "2R"), 4)
y1 <- rep(c("Fst.RsqAdjPos", "Fst.RsqPos", "Fst.Morans.I", "Fst.I.scaled"), each=3)
y2 <- rep(c("SES.RsqAdjPos", "SES.RsqPos", "SES.Morans.I", "SES.I.scaled"), each=3)
yLabels <- c(expression('Adjusted R'^2), "", "", expression('R'^2), "", "", 
             "Moran's I", "", "", "Scaled Moran's I", "", "")

for(i in 1:12)
{
  tmp1 <- tmp2 <- Results.table[[demo[i]]]
  tmp1 <- tmp1[ ,grep(y1[i], names(tmp1))]
  tmp2 <- tmp2[ ,grep(y2[i], names(tmp2))]
  names(tmp1) <- paste0("Y.", substr(names(tmp1), nchar(y1[i])+2, nchar(names(tmp1))))
  names(tmp2) <- paste0("Y.", substr(names(tmp2), nchar(y2[i])+2, nchar(names(tmp2))))
  
  rm(D11, D11.SES, D11.s, D11.SES.s)
  D11<- with(tmp1, data.frame(Comp.60=Y.drop.m / Y.90, 
                      Meta.60=Y.drop.MEM.m / Y.90, 
                      Meta.30=Y.30m / Y.90)) 
  D11.SES<- with(tmp2, data.frame(Comp.60=Y.drop.m / Y.90, 
                      Meta.60=Y.drop.MEM.m / Y.90, 
                      Meta.30=Y.30m / Y.90)) 
  D11<- gather(D11, "Comparison", "Y")
  D11.SES<- gather(D11.SES, "Comparison", "Y")
  D11.s <- groupwiseMean(Y ~ Comparison, data =D11, conf   = 0.95, digits = 3)
  D11.s$Comparison <- factor(D11.s$Comparison, levels = c("Comp.60", "Meta.60", "Meta.30"))
  D11.SES.s <- groupwiseMean(Y ~ Comparison, data =D11.SES, conf   = 0.95, digits = 3)
  D11.SES.s$Comparison <- factor(D11.SES.s$Comparison, levels = c("Comp.60", "Meta.60", "Meta.30"))
  #Plots.data[[i]] <- list(D11.s, D11.SES.s)
  
  Plots[[i]] <- ggplot(D11.s, aes(x = Comparison, y = Mean)) + ylim(-0.1,1.1) +
     geom_errorbar(aes(ymin = Trad.lower, ymax = Trad.upper), width = 0.05, size  = 0.5) +
     geom_point(shape = 1, size  = 2, colour="black") +
     theme_bw() + theme (panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
      ggtitle(demo[i]) + xlab("") + ylab(yLabels[i]) +
      theme(axis.text = element_text(size = 10),
            plot.title = element_text(size=16, face="bold", hjust = 0.5),
            axis.title.y = element_text(size=12, face="bold")) +
     geom_point(data = D11.SES.s, aes(x = Comparison, y = Mean), colour="grey70") +
     geom_errorbar(data = D11.SES.s, aes(ymin = Trad.lower, ymax = Trad.upper), 
                   width = 0.05, size  = 0.5, col="grey70") + 
     geom_hline(yintercept=1, size=0.5)
}

```

```{r fig.height=8, fig.width=6}

figure.3.P <- ggarrange(Plots[[1]], Plots[[2]], Plots[[3]], Plots[[4]], Plots[[5]],
                        Plots[[6]], Plots[[7]], Plots[[8]], Plots[[9]], Plots[[10]], 
                        Plots[[11]], Plots[[12]], ncol = 3, nrow = 4)
figure.3.P

g <- arrangeGrob(Plots[[1]], Plots[[2]], Plots[[3]], Plots[[4]], Plots[[5]],
                        Plots[[6]], Plots[[7]], Plots[[8]], Plots[[9]], Plots[[10]], 
                        Plots[[11]], Plots[[12]], ncol = 3, nrow = 4) 
ggsave(file=paste0(here::here(), "/output/Figure3.test.pdf"), g, width=8.5, height=12)
ggsave(file=paste0(here::here(), "/output/Figure3.test.jpg"), g, width=8.5, height=12)

```

With 9900 loci, no SES: NEED TO RUN RESULTS FOR DROP

```{r}
Results.table <- Results.subsets.table[[1]] %>% filter(NumInd == 20)
Results.table <- split(Results.table, Results.table$Demography)

#Plots.data <- list()
Plots <- list()

demo <- rep(c("IBD", "1R", "2R"), 3)
y1 <- rep(c("Fst.RsqAdjPos", "Fst.Morans.I", "Fst.I.scaled"), each=3)
y2 <- rep(c("SES.RsqAdjPos", "SES.Morans.I", "SES.I.scaled"), each=3)
yLabels <- c(expression('Adjusted R'^2), "", "", "Moran's I", "", "", 
                        "Scaled Moran's I", "", "")

for(i in 1:9)
{
  tmp1 <- tmp2 <- Results.table[[demo[i]]]
  tmp1 <- tmp1[ ,grep(y1[i], names(tmp1))]
  #tmp2 <- tmp2[ ,grep(y2[i], names(tmp2))]
  names(tmp1) <- paste0("Y.", substr(names(tmp1), nchar(y1[i])+2, nchar(names(tmp1))))
  #names(tmp2) <- paste0("Y.", substr(names(tmp2), nchar(y2[i])+2, nchar(names(tmp2))))
  
  rm(D11, D11.s)
  D11<- with(tmp1, data.frame(Comp.60=Y.drop.m / Y.90, 
                      Meta.60=Y.drop.MEM.m / Y.90, 
                      Meta.30=Y.30m / Y.90)) 

  D11<- gather(D11, "Comparison", "Y")
  D11.s <- groupwiseMean(Y ~ Comparison, data =D11, conf   = 0.95, digits = 3)
  D11.s$Comparison <- factor(D11.s$Comparison, levels = c("Comp.60", "Meta.60", "Meta.30"))

  #Plots.data[[i]] <- list(D11.s, D11.SES.s)
  
  Plots[[i]] <- ggplot(D11.s, aes(x = Comparison, y = Mean)) + ylim(0,1.1) +
     geom_errorbar(aes(ymin = Trad.lower, ymax = Trad.upper), width = 0.05, size  = 0.5) +
     geom_point(shape = 1, size  = 2, colour="black") +
     theme_bw() + theme (panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
      ggtitle(demo[i]) + xlab("") + ylab(yLabels[i]) +
      theme(axis.text = element_text(size = 10),
            plot.title = element_text(size=16, face="bold", hjust = 0.5),
            axis.title.y = element_text(size=12, face="bold"))  
}

figure.3.P <- ggarrange(Plots[[1]], Plots[[2]], Plots[[3]], Plots[[4]], Plots[[5]],
                        Plots[[6]], Plots[[7]], Plots[[8]], Plots[[9]], ncol = 3, nrow = 3)
figure.3.P

```
=======

## 8. Interaction Plots for supplementary material using the Dch as a measure for genetic distance

THIS NEEDS TO BE UPDATED!

### a) Figure S1

With n=90, #IND= 20 & the 9900 loci: Difference demographies in response variabe (Fst, adjusted R^2 and Moran's I).

```{r fig.height=2.5, fig.width=6}

#b <- which(Results.table$NumInd == 20)
#Results.table$Demography <- factor(Results.table$Demography, levels = c("IM", "IBD", "1R", "2R"))

Sum1.p <- groupwiseMean(Fst.90 ~ Demography, data = Results.table[b,], conf   = 0.95, digits = 3)
Pan.a.p<- ggplot(Sum1.p, aes(x = Demography, y = Mean)) + ylim(0.0,0.015) +
   geom_errorbar(aes(ymin = Trad.lower, ymax = Trad.upper), width = 0.05, size  = 0.5) +
   geom_point(shape = 1, size  = 2) +
   theme_bw() + theme (panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
    ggtitle("(a)") + xlab("Demography") + ylab("Fst") +
    theme(axis.text = element_text(size = 12),
          plot.title = element_text(size=16, face="bold"),
          axis.title.x = element_text(size=12, face="bold"),
          axis.title.y = element_text(size=12, face="bold")
          )

Sum2.p <- groupwiseMean(Dch.RsqAdjPos.90 ~ Demography, data = Results.table[b,], conf   = 0.95, digits = 3)
Pan.b.p<- ggplot(Sum2.p, aes(x = Demography, y = Mean)) + ylim(-0.02,0.8) +
   geom_errorbar(aes(ymin = Trad.lower, ymax = Trad.upper), width = 0.05, size  = 0.5) +
   geom_point(shape = 1, size  = 2) +
   theme_bw() + theme (panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
    ggtitle("(b)") + xlab("Demography") + ylab(expression(bold('Adjusted R'^2))) +
    theme(axis.text = element_text(size = 12),
          plot.title = element_text(size=16, face="bold"),
          axis.title.x = element_text(size=12, face="bold"),
          axis.title.y = element_text(size=12, face="bold")
          )

Sum3.p <- groupwiseMean(Dch.Morans.I.90 ~ Demography, data = Results.table[b,], conf   = 0.95, digits = 3)
Pan.c.p<- ggplot(Sum3.p, aes(x = Demography, y = Mean)) + ylim(-0.02,0.8) +
   geom_errorbar(aes(ymin = Trad.lower, ymax = Trad.upper), width = 0.05, size  = 0.5) +
   geom_point(shape = 1, size  = 2) +
   theme_bw() + theme (panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
    ggtitle("(c)") + xlab("Demography") + ylab("Moran's I") +
    theme(axis.text = element_text(size = 12),
          plot.title = element_text(size=16, face="bold"),
          axis.title.x = element_text(size=12, face="bold"),
          axis.title.y = element_text(size=12, face="bold")
          )
figure.1.P <- ggarrange(Pan.a.p, Pan.b.p, Pan.c.p, ncol = 3, nrow = 1)
#figure.1.P

g <- arrangeGrob(Pan.a.p, Pan.b.p, Pan.c.p, ncol = 3, nrow = 1) #generates g
 ggsave(file=paste0(here::here(), "/output/FigureS1.pdf"), g)

```

### b) Figure S2

With n=90 and removing IM: Interaction between #Ind (20, 6) and # Loci (9900, 3300, 500) in response variabe (Fst, adjusted R^2, and Moran's I).


```{r fig.height=2.5, fig.width=6}

loci <- c(9900, 3300, 500, 100, 50)

s = 3

Results.sub.wide <- Results.table[[s]] %>% 
  mutate(Demography = as.character(Demography), NumInd = as.character(NumInd)) %>%
  select(Demography, NumPops, NumInd, Fst.90, Fst.90.3, Fst.90.500, Dch.RsqAdjPos.90, Dch.Morans.I.90, 
         Dch.I.scaled.90, Dch.RsqAdjPos.90.3, Dch.Morans.I.90.3, Dch.I.scaled.90.3, 
         Dch.RsqAdjPos.90.500, Dch.Morans.I.90.500, Dch.I.scaled.90.500)

# Convert to long format
Results.sub.long <- data.frame(
  Results.sub.wide %>% 
    select(Demography, NumPops, NumInd, Fst.90, Fst.90.3, Fst.90.500) %>%
    melt(id.vars=c("Demography", "NumPops", "NumInd")) %>% rename(Fst = value) %>% 
    select(Demography, NumPops, NumInd, Fst),
  Results.sub.wide %>% 
    select(Demography, NumPops, NumInd, Dch.RsqAdjPos.90, Dch.RsqAdjPos.90.3, Dch.RsqAdjPos.90.500) %>%
    melt(id.vars=c("Demography", "NumPops", "NumInd")) %>% rename(Adj.Rsqr.Pos = value) %>% 
    select(Adj.Rsqr.Pos),
  Results.sub.wide %>% 
    select(Demography, NumPops, NumInd, Dch.Morans.I.90, Dch.Morans.I.90.3, Dch.Morans.I.90.500) %>%
    melt(id.vars=c("Demography", "NumPops", "NumInd")) %>% rename(Morans.I = value) %>% 
    select(Morans.I),
  Results.sub.wide %>% 
    select(Demography, NumPops, NumInd, Dch.I.scaled.90, Dch.I.scaled.90.3, Dch.I.scaled.90.500) %>%
    melt(id.vars=c("Demography", "NumPops", "NumInd")) %>% rename(I.scaled = value) %>% 
    select(I.scaled))

# Add number of loci, drop IM
Results.sub.long <- Results.sub.long %>% mutate(NumLoci = rep(c(9900, 3300, 500), each=24)) %>% 
  mutate(NumLoci = factor(NumLoci, levels = c("500", "3300", "9900"))) %>% 
  filter(Demography != "IM")

pdf(file = paste0(here::here(), "/output/FigureS2.pdf"), width=8, height=3.5)

par(mfcol=c(1,3), mar = c(5, 5, 3, 1))
  with(Results.sub.long, {interaction.plot(NumInd, NumLoci, Fst, 
                                             col=c("grey1","grey20","grey50"), 
                                             ylab=" mean Fst", ylim=c(0.0,0.015), lwd= 3, 
                                             cex.lab=1.5, cex.axis=1.5, cex.sub=1.5, 
                                             xleg = "bottomright")})
    title("(a)", adj = 0, line =1, cex.main=2 )
    
  with(Results.sub.long, {interaction.plot(NumInd, NumLoci, Adj.Rsqr.Pos, 
                                                  col=c("grey1","grey20","grey50"), 
                                                  ylab=expression("mean adjusted R"^2), ylim=c(0.0,0.6), 
                                                  lwd= 3, cex.lab=1.5, cex.axis=1.5, cex.sub=1.5, 
                                                  legend=FALSE)})
  title("(b)", adj = 0, line =1, cex.main=2 )
  
  with(Results.sub.long, {interaction.plot(NumInd, NumLoci, Morans.I, 
                                                  col=c("grey1","grey20","grey50"), 
                                                  ylab="mean Moran's I", ylim=c(0.0,0.6), 
                                                  lwd= 3, cex.lab=1.5, cex.axis=1.5, cex.sub=1.5, 
                                                  legend=FALSE)})
  title("(c)", adj = 0, line =1, cex.main=2 )

dev.off()

```

### c) Figure S3 

Multipanel fig. one figure for each demography all on the same scale. boxplots (20 ind & 9900 loci)

Respo= Adj R^2 & MI & scaled MI. (x axis = 60 refitting, missing values, 30 refitting)... show variability as proportion (divide each value by the value of the 90)

```{r fig.height=6, fig.width=6}

# Dch Adjusted Rsq.

D11<- with(Results.table %>% filter(Demography == "IBD" & NumInd == 20), 
           data.frame(Comp.60=Dch.RsqAdjPos.drop.m / Dch.RsqAdjPos.90, 
                      Meta.60=Dch.RsqAdjPos.drop.MEM.m / Dch.RsqAdjPos.90, 
                      Meta.30=Dch.RsqAdjPos.30.m / Dch.RsqAdjPos.90)) 
D11<- gather(D11, "Comparison", "Dch.Rsq")
D11.s <- groupwiseMean(Dch.Rsq ~ Comparison, data =D11, conf   = 0.95, digits = 3)
D11.s$Comparison <- factor(D11.s$Comparison, levels = c("Comp.60", "Meta.60", "Meta.30"))
D11.a.s<- ggplot(D11.s, aes(x = Comparison, y = Mean)) + ylim(0,1.0) +
   geom_errorbar(aes(ymin = Trad.lower, ymax = Trad.upper), width = 0.05, size  = 0.5) +
   geom_point(shape = 1, size  = 2) +
   theme_bw() + theme (panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
    ggtitle("IBD") + xlab("") + ylab(expression(bold('Adjusted R'^2))) +
    theme(axis.text = element_text(size = 10),
          plot.title = element_text(size=16, face="bold", hjust = 0.5),
          axis.title.y = element_text(size=12, face="bold")
          )

D22<- with(Results.table %>% filter(Demography == "1R" & NumInd == 20), 
           data.frame(Comp.60=Dch.RsqAdjPos.drop.m / Dch.RsqAdjPos.90, 
                      Meta.60=Dch.RsqAdjPos.drop.MEM.m / Dch.RsqAdjPos.90, 
                      Meta.30=Dch.RsqAdjPos.30.m / Dch.RsqAdjPos.90)) 
D22<- gather(D22, "Comparison", "Dch.Rsq")
D22.s <- groupwiseMean(Dch.Rsq ~ Comparison, data =D22, conf   = 0.95, digits = 3)
D22.s$Comparison <- factor(D22.s$Comparison, levels = c("Comp.60", "Meta.60", "Meta.30"))
D22.a.s<- ggplot(D22.s, aes(x = Comparison, y = Mean)) + ylim(0,1.0) +
   geom_errorbar(aes(ymin = Trad.lower, ymax = Trad.upper), width = 0.05, size  = 0.5) +
   geom_point(shape = 1, size  = 2) +
   theme_bw() + theme (panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
    ggtitle("1R") + xlab("") + ylab("") +
    theme(axis.text = element_text(size = 10),
          plot.title = element_text(size=16, face="bold", hjust = 0.5),
          axis.title.y = element_text(size=12, face="bold")
          )

D33<- with(Results.table %>% filter(Demography == "2R" & NumInd == 20), 
           data.frame(Comp.60=Dch.RsqAdjPos.drop.m / Dch.RsqAdjPos.90, 
                      Meta.60=Dch.RsqAdjPos.drop.MEM.m / Dch.RsqAdjPos.90, 
                      Meta.30=Dch.RsqAdjPos.30.m / Dch.RsqAdjPos.90)) 
D33<- gather(D33, "Comparison", "Dch.Rsq")
D33.s <- groupwiseMean(Dch.Rsq ~ Comparison, data =D33, conf   = 0.95, digits = 3)
D33.s$Comparison <- factor(D33.s$Comparison, levels = c("Comp.60", "Meta.60", "Meta.30"))
D33.a.s<- ggplot(D33.s, aes(x = Comparison, y = Mean)) + ylim(0,1.0) +
   geom_errorbar(aes(ymin = Trad.lower, ymax = Trad.upper), width = 0.05, size  = 0.5) +
   geom_point(shape = 1, size  = 2) +
   theme_bw() + theme (panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
    ggtitle("2R") + xlab("") + ylab("") +
    theme(axis.text = element_text(size = 10),
          plot.title = element_text(size=16, face="bold", hjust = 0.5),
          axis.title.y = element_text(size=12, face="bold")
          )


D44<- with(Results.table %>% filter(Demography == "IBD" & NumInd == 20),
           data.frame(Comp.60=Dch.Morans.I.drop.m / Dch.Morans.I.90, 
                      Meta.60=Dch.Morans.I.drop.MEM.m / Dch.Morans.I.90, 
                      Meta.30=Dch.Morans.I.30.m / Dch.Morans.I.90)) 
D44<- gather(D44, "Comparison", "Dch.Rsq")
D44.s <- groupwiseMean(Dch.Rsq ~ Comparison, data =D44, conf   = 0.95, digits = 3)
D44.s$Comparison <- factor(D44.s$Comparison, levels = c("Comp.60", "Meta.60", "Meta.30"))
D44.a.s<- ggplot(D44.s, aes(x = Comparison, y = Mean)) + ylim(0,1.0) +
   geom_errorbar(aes(ymin = Trad.lower, ymax = Trad.upper), width = 0.05, size  = 0.5) +
   geom_point(shape = 1, size  = 2) +
   theme_bw() + theme (panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
   xlab("") + ylab("Moran's I") +
    theme(axis.text = element_text(size = 10),
          axis.title.y = element_text(size=12, face="bold")
          )

D55<- with(Results.table %>% filter(Demography == "1R" & NumInd == 20), 
           data.frame(Comp.60=Dch.Morans.I.drop.m / Dch.Morans.I.90, 
                      Meta.60=Dch.Morans.I.drop.MEM.m / Dch.Morans.I.90, 
                      Meta.30=Dch.Morans.I.30.m / Dch.Morans.I.90)) 
D55<- gather(D55, "Comparison", "Dch.Rsq")
D55.s <- groupwiseMean(Dch.Rsq ~ Comparison, data =D55, conf   = 0.95, digits = 3)
D55.s$Comparison <- factor(D55.s$Comparison, levels = c("Comp.60", "Meta.60", "Meta.30"))
D55.a.s<- ggplot(D55.s, aes(x = Comparison, y = Mean)) + ylim(0,1.0) +
   geom_errorbar(aes(ymin = Trad.lower, ymax = Trad.upper), width = 0.05, size  = 0.5) +
   geom_point(shape = 1, size  = 2) +
   theme_bw() + theme (panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
   xlab("") + ylab("") +
    theme(axis.text = element_text(size = 10),
          axis.title.y = element_text(size=12, face="bold")
          )

D66<- with(Results.table %>% filter(Demography == "2R" & NumInd == 20), 
           data.frame(Comp.60=Dch.Morans.I.drop.m / Dch.Morans.I.90, 
                      Meta.60=Dch.Morans.I.drop.MEM.m / Dch.Morans.I.90, 
                      Meta.30=Dch.Morans.I.30.m / Dch.Morans.I.90)) 
D66<- gather(D66, "Comparison", "Dch.Rsq")
D66.s <- groupwiseMean(Dch.Rsq ~ Comparison, data =D66, conf   = 0.95, digits = 3)
D66.s$Comparison <- factor(D66.s$Comparison, levels = c("Comp.60", "Meta.60", "Meta.30"))
D66.a.s<- ggplot(D66.s, aes(x = Comparison, y = Mean)) + ylim(0,1.0) +
   geom_errorbar(aes(ymin = Trad.lower, ymax = Trad.upper), width = 0.05, size  = 0.5) +
   geom_point(shape = 1, size  = 2) +
   theme_bw() + theme (panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
   xlab("") + ylab("") +
    theme(axis.text = element_text(size = 10),
          axis.title.y = element_text(size=12, face="bold")
          )

# Scaled Moran's I

D77<- with(Results.table %>% filter(Demography == "IBD" & NumInd == 20), 
           data.frame(Comp.60=Dch.I.scaled.drop.m / Dch.I.scaled.90,
                      Meta.60=Dch.I.scaled.drop.MEM.m / Dch.I.scaled.90, 
                      Meta.30=Dch.I.scaled.30.m / Dch.I.scaled.90)) 
D77<- gather(D77, "Comparison", "Dch.Rsq")
D77.s <- groupwiseMean(Dch.Rsq ~ Comparison, data =D77, conf   = 0.95, digits = 3)
D77.s$Comparison <- factor(D77.s$Comparison, levels = c("Comp.60", "Meta.60", "Meta.30"))
D77.a.s<- ggplot(D77.s, aes(x = Comparison, y = Mean)) + ylim(0,1.0) +
   geom_errorbar(aes(ymin = Trad.lower, ymax = Trad.upper), width = 0.05, size  = 0.5) +
   geom_point(shape = 1, size  = 2) +
   theme_bw() + theme (panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
   xlab("") + ylab("Scaled Moran's I") +
    theme(axis.text = element_text(size = 10),
          axis.title.y = element_text(size=12, face="bold")
          )

D88<- with(Results.table %>% filter(Demography == "1R" & NumInd == 20), 
           data.frame(Comp.60=Dch.I.scaled.drop.m / Dch.I.scaled.90,  
                      Meta.60=Dch.I.scaled.drop.MEM.m / Dch.I.scaled.90, 
                      Meta.30=Dch.I.scaled.30.m / Dch.I.scaled.90)) 
D88<- gather(D88, "Comparison", "Dch.Rsq")
D88.s <- groupwiseMean(Dch.Rsq ~ Comparison, data =D88, conf   = 0.95, digits = 3)
D88.s$Comparison <- factor(D88.s$Comparison, levels = c("Comp.60", "Meta.60", "Meta.30"))
D88.a.s<- ggplot(D88.s, aes(x = Comparison, y = Mean)) + ylim(0,1.0) +
   geom_errorbar(aes(ymin = Trad.lower, ymax = Trad.upper), width = 0.05, size  = 0.5) +
   geom_point(shape = 1, size  = 2) +
   theme_bw() + theme (panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
   xlab("") + ylab("") +
    theme(axis.text = element_text(size = 10),
          axis.title.y = element_text(size=12, face="bold")
          )

D99<- with(Results.table %>% filter(Demography == "2R" & NumInd == 20), 
           data.frame(Comp.60=Dch.I.scaled.drop.m / Dch.I.scaled.90,  
                      Meta.60=Dch.I.scaled.drop.MEM.m / Dch.I.scaled.90, 
                      Meta.30=Dch.I.scaled.30.m / Dch.I.scaled.90)) 
D99<- gather(D99, "Comparison", "Dch.Rsq")
D99.s <- groupwiseMean(Dch.Rsq ~ Comparison, data =D99, conf   = 0.95, digits = 3)
D99.s$Comparison <- factor(D99.s$Comparison, levels = c("Comp.60", "Meta.60", "Meta.30"))
D99.a.s<- ggplot(D99.s, aes(x = Comparison, y = Mean)) + ylim(0,1.02) +
   geom_errorbar(aes(ymin = Trad.lower, ymax = Trad.upper), width = 0.05, size  = 0.5) +
   geom_point(shape = 1, size  = 2) +
   theme_bw() + theme (panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
    xlab("") + ylab("") +
    theme(axis.text = element_text(size = 10),
          axis.title.y = element_text(size=12, face="bold")
          )

figure.3.P <- ggarrange(D11.a.s, D22.a.s, D33.a.s, D44.a.s, D55.a.s, D66.a.s ,D77.a.s, D88.a.s, D99.a.s, ncol = 3, nrow = 3)
#figure.3.P

g <- arrangeGrob(D11.a.s, D22.a.s, D33.a.s, D44.a.s, D55.a.s, D66.a.s ,D77.a.s, D88.a.s, D99.a.s, ncol = 3, nrow = 3) 
ggsave(file=paste0(here::here(), "/output/FigureS3.pdf"), g)

```

